Light inextensible strings (thread) under tension in the Schwarzschild geometry

### 2021-02-27 tpe_edreview_rebuttal.txt

Fixed, notation now consistent

> 
> 11.  The paragraph below Eq. (6) is puzzling.  Here the path of the
> tight string is being compared to an "equatorial null geodesic," which
> I take to mean the path of a light beam moving along the equator.  If
> we are considering a light beam, then shouldn't a term with $t$ in it
> appear somewhere?  The puzzlement continues given the statement that
> "light, being sensitive to the time component of the metric" whereas
> the equations discussed in this paragraph say nothing (that I can see)
> about the time component of the metric.  The proper time interval
> would be zero, but the coordinate time interval $dt$ would, I think,
> be nonzero.  What am I missing?
>

The equation given as $u''=3u^2/2-u$ gives the path of a beam of light
through Schwarzschild space ideally in terms of $r$ as a function of
angle, $r=r(\phi)$.  The idea is that a stationary observer has a
laser pointer pointing in a particular direction, switches it on and
leaves it on for a long time.  We then record all the points that the
beam touches.  The path of the light is then simply $r=r(\phi)$, no
time coordinate involved, as Figures 4 and 6 show.  I have changed the
citation to Rindler, who gives a detailed derivation.

> 12.  In the passage below Eq. (6), I am having trouble making the
> connection between $r''(\phi)=r-3M + 2(r')^2/r$ and the claim that the
> solution is $r=r_0\sec(\phi-\phi_0)$.  What trig identity am I
> missing?
>

There are two separate issues here.  Firstly, null geodesic paths in
the Schwarzschild metric obey $u''=3u^2/2-u$ where $u=2M/r$.  This is
equivalent to $r''(\phi) = r-3M + 2(r')^2/r$.  The second issue is
straight lines in Euclidean flat space (in flat space the expression
"straight lne" is unambiguous).  WLOG $\phi_0=0$ and if
$r=r_0\sec\phi$ then $r'=-r_0\sin\phi\cdot\sec^2\phi$ and
$r''=r_0(\sec\phi + 2\sin^2\phi\sec^3\phi$, from which $r''=2(r')^2/r$
follows from cancellation.

One can "go the other way"---that is, given ODE $r''=r+2(r')^2/r$, one
can deriving $r=r_0\sec\phi$---by substituting $u=1/r$ and obtaining
$u''+u=0$, whence $u=A\cos(\phi-\phi_0)$.

I have rephrased slightly and numbered one of the equations so I can
refer to it.

> 13.  Can the author explain more fully what is meant by a "tangential
> null geodesic" (several lines after Eq. 6)?  That terminology makes it
> sound like the r-coordinate does not change, whence $r'$ would be
> zero, but that appears to not be the case.  In this section, if $r'$
> is small I see where the approximation $r''=r-2M$ comes from, but I am
> having trouble seeing where $r''-r-3M$ arises for the alleged
> tangential null geodesics.  What am I not seeing?

> 
> I will have to take the author's word for the numerical results.
> 
> This is an interesting paper.  I congratulate the author for thinking
> of such an intriguing application of the Schwarzschild metric.
> However, in this reviewer's opinion, before the paper can be well
> fitted to its announced audience ("an undergraduate homework problem")
> some points need to be clarified.

### 2021-02-27 string_physics_educator_revision3.tex
fixes #45

\noindent Solutions to equation~\ref{rprimeprime} correspond to a taut
string confined to Flamm's paraboloid~\cite{flamm1916}; see
Figure~\ref{flamm}.  It is interesting to compare
equation~\ref{rprimeprime} with the corresponding equation for
equatorial null geodesics, which is usually given~\cite{wald} in terms
of $u=2M/r$ as $u''=3u^2/2-u$, but is equivalent
to~$r''\left(\phi\right)=r-3M+2\left(r'\right)^2/r$.  This shows that
light, being sensitive to the time component of the metric, behaves
differently from string, which is not (at this point it might be worth
``reminding'' students that a straight line in flat space,
$r=r_0\sec\left(\phi-\phi_0\right)$, obeys
$r''=r+2\left(r'\right)^2/r$).

>>>>

\noindent Solutions to equation~\ref{rprimeprime} correspond to a taut
string confined to Flamm's paraboloid~\cite{flamm1916}; see
Figure~\ref{flamm}.  It is interesting to compare
equation~\ref{rprimeprime} with the corresponding equation for
equatorial null geodesics, which is usually given~\cite{rindler} in
terms of $u=2M/r$ as $u''=3u^2/2-u$, but is equivalent
to

\begin{equation}\label{nullgeodesicODE}
  r''\left(\phi\right)=r-3M+2\left(r'\right)^2/r.
\end{equation}

This shows that
light, being sensitive to the time component of the metric, behaves
differently from string, which is not.  At this point it might be worth
observing that a straight line in flat space,
$r=r_0\sec\left(\phi-\phi_0\right)$, obeys
$r''=r+2\left(r'\right)^2/r$: identical to equation~\ref{nullgeodesicODE} but without the $M$ term.
